---
title: "MappingIL4"
format: html
editor: visual
---

## Illinoi dsistrict 4

To start with, load the packages we need.

```{r chunk01, results='hide'}
library(sf)
library(dplyr)
library(ggplot2)
```

Define the common directory where we'll save local copies of the things we might want to use in other notebooks.

```{r chunk02}
my_data_dir <- rappdirs::user_data_dir(appauthor = "HarshaYerneni", appname = "Mappingil4")
if (!dir.exists(my_data_dir)) dir.create(my_data_dir, recursive = TRUE)
tools::file_path_as_absolute(my_data_dir)

```

Load the shapefile with the Illinois Congressional district boundaries that we downloaded from redistrictingdatahub.org

```{r chunk03, cache = TRUE}
il_congdst_sf <- sf::st_read("il_cong_adopted_2021/HB 1291 FA #1.shp")

nrow(il_congdst_sf)
head(il_congdst_sf)
```

Given that we know that Illinois currently has 17 districts, this appears to be consistent.

Let's plot the district boundaries using the geometries we just loaded from the shapefile

```{r chunk04, cache = TRUE}
ggplot(il_congdst_sf) + 
  geom_sf() + 
  labs(title = "Illinois Congressional Districts, 2021")

```

Let's pull the IL-4 district geometries

```{r chunk05, cache = TRUE}
il4 <- dplyr::filter(il_congdst_sf, DISTRICT == 04)

```

Let's plot IL-4

```{r chunk06, cache = TRUE}
ggplot(il4) + 
  geom_sf() + 
  labs(title = "Illinois Congressional District 4")

```

We are interested in the precinct boundaries.

```{r chunk07, cache = TRUE}
il_prec_2022_sf <- sf::st_read("il_2022_gen_prec/il_2022_gen_cong_prec/il_2022_gen_cong_prec.shp")

nrow(il_prec_2022_sf)
head(il_prec_2022_sf)

```

Let's look at all the column names in this table to see what's in the table

```{r chunk08, cache = TRUE}

names(il_prec_2022_sf)

```

County County Name Precinct Precinct Name CONG_DIST Congressional District

```{r chunk09, cache = TRUE}
il4_prec <- dplyr::filter(il_prec_2022_sf, CONG_DIST == 04)
head(il4_prec)
```

```{r chunk10, cache = TRUE}
il4_prec_sel <- il4_prec %>% select(UNIQUE_ID,
                                      COUNTYFP,
                                      County,
                                      Precinct,
                                      CONG_DIST,
                                      GCON04DGAR, 
                                      GCON04OWRI, 
                                      GCON04RFAL, 
                          
                                      geometry)
head(il4_prec_sel)

```

```{r chunk11, cache = TRUE}
ggplot(il4_prec_sel) + 
  geom_sf() + 
  geom_sf(data = il4, fill = NA, lwd = 1.2) +
  labs(title = "Precincts in Illinois Congressional District 4")

```

This map does appear to have some precincts that extend outside of the IL-4 Congressional district boundary. This needs further investigation.

```{r chunk12, cache = TRUE}
il4_prec_win <- il4_prec_sel %>% mutate(
  Winner = case_when (
    GCON04DGAR > GCON04RFAL ~ "D",
    GCON04DGAR < GCON04RFAL ~ "R",
    TRUE ~ "T") )

head(il4_prec_win)

```

We have a few ties. One with no votes that might need investigation and another with each side getting 125 votes.

Lets maps the winners

```{r chunk13, cache = TRUE}
ggplot(il4_prec_win, aes(color = Winner)) + 
  geom_sf() + 
  scale_color_manual(values = c("D" = "blue",
                                "R" = "red",
                                "T" = "green"))
  


```

```{r chunk14, cache = TRUE}
ggplot(il4_prec_win, aes(fill = Winner)) + 
  geom_sf() 

```

```{r chunk15, cache = TRUE}
ggplot(il4_prec_win, aes(fill = Winner)) + 
  geom_sf() + 
  scale_fill_manual(values = c("D" = "blue",
                                "R" = "red",
                                "T" = "green"))


```

```{r chunk16, cache = TRUE}
ggplot(il4_prec_win, aes(fill = Winner, color = "white")) + 
  geom_sf() + 
  scale_fill_manual(values = c("D" = "blue",
                                "R" = "red",
                                "T" = "green"))
  

```

```{r chunk16, cache = TRUE}
ggplot(il4_prec_win, aes(fill = Winner)) + 
  geom_sf() + 
  scale_fill_manual(values = c("D" = "blue",
                                "R" = "red",
                                "T" = "green"))+
  geom_sf(data = il4, fill = NA, lwd = 0.8) +
  labs(title = "2022 IL Cong District 4 Election Results by Precinct")
```

```{r chunk17 cache=TRUE}
totalDvotes <- sum(il4_prec_win$GCON04DGAR)
totalRvotes <- sum(il4_prec_win$GCON04RFAL)
totalOvotes <- sum(il4_prec_win$GCON04OWRI)
cat(totalDvotes, "votes for the Democrat\n")
cat(totalRvotes, "votes for the Republican\n")
cat(totalOvotes, "votes for others\n")

```

```{r chunk 18}
library(readr)
```

```{r chunk19}
# Load ing libraries
library(tidycensus)
library(sf)
library(ggplot2)
library(dplyr)
#census_api_key("41d067309a9134ca496c16adfd33605fb13d1cb4", install = TRUE)
# Function to load and process district data
load_district_data <- function(variable, year = 2020) {
  # Retrieve data for Illinois congressional districts
  district_data <- get_acs(
    geography = "congressional district",
    variables = variable,
    state = "IL",
    year = year,
    geometry = TRUE
  )
  # Check structure to confirm column names
  print(head(district_data))
  # Filter for District 4 using GEOID if NAME column is not present
  district_4 <- district_data %>%
    filter(GEOID == "1704") %>%
    st_transform(crs = 4269)  # Transform to NAD 83
  return(district_4)
}
# Function to plot a specific feature of District 4
plot_district_feature <- function(gdf, feature) {
  ggplot(data = gdf) +
    geom_sf(aes_string(fill = feature)) +
    scale_fill_viridis_c() +
    ggtitle(paste("Illinois District 4 -", feature)) +
    theme_minimal()
}
variable <- "B01003_001E"  # Total population variable for ACS
district_data <- load_district_data(variable)
# Plot the selected feature
plot_district_feature(district_data, "estimate")

```

```{r chunk20}
# Function to load and process district data for a specific ethnicity variable
load_district_ethnicity_data <- function(variable, year = 2020) {
  # Retrieve data for Illinois congressional districts
  district_data <- get_acs(
    geography = "congressional district",
    variables = variable,
    state = "IL",
    year = year,
    geometry = TRUE
  )
  # Filter for District 4 using GEOID if NAME column is not present
  district_4 <- district_data %>%
    filter(GEOID == "1704") %>%
    st_transform(crs = 4269)  # Transform to NAD 83
  return(district_4)
}
# Function to plot ethnicity for District 4
plot_district_ethnicity <- function(gdf, feature) {
  ggplot(data = gdf) +
    geom_sf(aes_string(fill = feature)) +
    scale_fill_viridis_c() +
    ggtitle(paste("Illinois District 4 -", feature, "(Hispanic or Latino Population)")) +
    theme_minimal()
}

# Example usage
variable <- "B03002_012E"  # Hispanic or Latino population variable
district_ethnicity_data <- load_district_ethnicity_data(variable)
# Plot the Hispanic or Latino population in Illinois District 4
plot_district_ethnicity(district_ethnicity_data, "estimate")

```

```{}
```

```{}
```

```{}
```

```{}
```

```{}
```

```{}
```

```{}
```
